Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2018-08-14 Ben_Manuscripts/structure_paper/Draft.tex
Comments.

  We further refined our structural understanding of the system by simulating X-ray 
  diffraction patterns produced from equilibrated MD trajectories and comparing them
  to experiment. We tested systems built with 5 columns per pore in the parallel 
  displaced and sandwiched configurations at 300 K in the ordered and disordered
  basins. We generated simulated patterns using the equilibrated portion of each 
  simulation trajectory. The patterns for all structures are shown and compared to
  experiment in Figure~\ref{fig:XRDsim}.

  The diffraction patterns show some noise, especially along the axis where $q_r$=0. There are
  two main reasons for this noise. First is that we angle averaged the 3D structure factor 
  (see section~\ref{method:xrd}) which inherently gives less data as $q_r$ approaches 0. Second, 
  our simulations are not long enough to sample enough truly independent configurations within
  their respective metastable basins. By sampling more uncorrelated equilibrium configurations,
  we can suppress some of the noise (See section S-\ref{S-section:xrd_noise}).

@@

  The simulated XRD patterns show moderate qualitative agreement with experiment. R-alkanes and
  R-pores appear in the expected location. R-pores is more intense than experiment, likely
  because we are simulating a near perfect, infinite hexagonal array. The real system has 
  defects and domain misalignment which decreases the overall intensity.
  % MRS4: comments from Matt suggest that domain misalignment shouldn't alone be responsible for 
  % the difference in intensity ratio. Some of the experiments should help explain it better. 	
  % Let's talk!
  % BJC4: So a combo of: domain misalignment, defects, slow dynamics of simulated system, only 4 
  % pores in unit cell, tortuosity (?)
  % MRS5: maybe, there's a question of how we can demonstrate the magnitude of any of these 
  % effects.
  R-spots appears to be weaker, except for the ordered basin sandwiched configuration, in the 
  simulated patterns. R-spots is also partially engulfed by the wide R-$\pi$ reflection. R-$\pi$
  appears at a lower $q_z$ value than experiment. In all systems, the reflection reaches its 
  maximum at $q_z < 1.5~\AA^{-1}$ which means that monomers prefer to stack at least 4.2 \AA~apart
  rather than 3.7 \AA. R-double does not appear in any of the patterns.

@@

  We quantified the numerical discrepancies present when comparing the relative intensities
  of each reflection of interest between experimental and simulated patterns. 
  Table~\ref{table:relative_inensities_300K} shows the relative intensities of each reflection for 
  all systems tested. The patterns are normalized so that the average intensity of R-alkanes 
  must equal 1. We measured the approximate intensity of R-$\pi$ and R-double by measuring the
  intensity of the appropriate peaks of the cross-section of the 2D pattern at $q_r=0$. The 
  relative intensity of R-$\pi$ is significantly higher than experiment in our simulations. 
  R-spots is measured as the average intensity within the region bounded by a 'spot'. We 
  identified spots based on visual inspection. If the spots were not easily discernible, then the 
  intensity was taken as that of the intersection of R-alkanes at half the $q_z$ value of
  R-$\pi$, since that is where it appears experimentally. The intensity of R-spots is slightly
  lower than experiment in all cases except for the ordered basin sandwiched configuration which 
  agrees with experiment. There is no R-double intensity to be measured. Figures
  are provided in section S-\ref{S-section:xrd_intensities} to further clarify these measurements.

  \begin{table}[h]
  \centering
  \begin{tabular}{c|ccccc}
  \toprule
 		   & \multicolumn{5}{c}{Configuration} \\
  \hline
             &            &            & Parallel  & Disordered & Disordered         \\
  Reflection & Experiment & Sandwiched & Displaced & Sandwiched & Parallel Displaced \\
  \midrule
  R-alkanes  & 1.0        &  1.0       &  1.0      & 1.0        & 1.0                \\  
  R-spots    & 1.3        &  1.3       &  1.2      & 1.1        & 1.2                \\
  R-$\pi$    & 2.8        & 44.0       &  7.7      & 8.4        & 10.1               \\
  R-double   & 0.9        &  --        & --        &  --        & --                 \\ 
  \bottomrule
  \end{tabular}
  \caption{The simulated XRD patterns of the systems tested, normalized so that the average 
  intensity of R-alkanes equals 1, show R-$\pi$ reflections that are significantly higher 
  than experiment and R-spots reflections that are slightly lower than experiment. R-double
  does not appear in any patterns, and thus has no measurable intensity.}
  \label{table:relative_inensities_300K} 
  \end{table}
  %BJC5: Bold font is caused by achemso document type
  %MRS7: huh.
  %BJC6: yea, specifically the one for j phys chem B. I compiled with a different document type and the captions looked normal
  There are clear differences between simulated and experimental results that must be addressed
  and justified. We will explore the possible origin of select experimental reflections and form 
  structural hypotheses related to their appearance in our simulated patterns. Specifically, we 
  want answer:  
  \begin{enumerate}
	\item What is the origin of R-spots?
  	\item Why is the intensity of R-$\pi$ so much greater than experiment?
  	\item What is the origin of R-double?
  \end{enumerate}

@@

  \subsubsection{The origin of R-spots}\label{section:rspots}

  In order to more clearly study R-spots, we equilibrated an ordered basin sandwiched 
  configuration at 280K. Obtaining an equilibrium structure at a precise temperature is
  a challenge that force fields often fail to overcome. Wang et al. highlighted the 
  shortcoming of some of the most popular protein force field in predicting the temperature
  dependence of protein structural ensembles \cite{wang_building_2017}. It is possible that   % page 12
  our membrane system, simulated at 300K, does not fully represent the true structure at 300K.
  Lowering the temperature to 280K helped us to better resolve R-spots (Figure~\ref{fig:sandwiched280K}). 
%  The relative intensity 
%  of R-spots in this system is slightly higher than experiment (1.5 vs. 1.3)

@@

  Evidence strongly suggests that the R-spots signal is not a result of alkane
  chain tilt. Previous literature has attributed the spots in this particular
  WAXS pattern as the product of tilted alkane chains~\cite{feng_scalable_2014}.
  We measured the tilt angle of the alkane chains of the 280K system and showed
  that it equilibrates to an average tilt angle of
  -2\degree~(Fig.~\ref{fig:tilt}), far from the 37\degree~tilt angle previously
  used to explain R-spots. 
  %MRS8: did you do a few different things here as sensitivity analysis to show that the finding was robust?  If so, putting some of that in would be useful.
  %BJC7: make it clear that it is the 280K. Reference all configurations where tilt angles were calculated.

@@

  %BJC9: I think Matt gets confused here. I put a more detailed explanation in 
  % the supplemental information. Do you think that's enough, or should I add
  % to this figure?
  The tails pack together hexagonally. We plotted the center of masses of the 
  first four tail atoms of each tail originating from each head group 
  (Figure \ref{fig:centroids}). We measured the angle between each center of
  mass and its nearest neighbor center of masses with respect to the $xy$ 
  plane of the membrane. We see distinct peaks in the distribution of these 
  angles located ca. $-60\degree$, $0\degree$, and $60\degree$ which is 
  consistent with a hexagonally packed configuration 
  (Figure~\ref{fig:layered_tails}). This ordering dies off at the tail ends,
  where there is more space to fill causing them to pack nearly isotropically.
  See Section~S-\ref{S-tail_packing} for a more detailed explanation of the 
  calculation and packing distributions generated from different sections of
  the tails.

  The peaks in the nearest neighbor angle distribution are consistent with the
  location of R-spots. The 2D Fourier transform of a simple hexagonal array 
  constructed based on the peak angles in Figure~\ref{fig:layered_tails}
  shows reflections in the same locations as R-spots, in addition to vertical 
  stacking reflections along the $q_z$ axis that would intersect with R-$\pi$
  (Figure~\ref{fig:hexagonal_ft}).

@@

  \subsubsection{Discrepancies between experimental and simulated R-$\pi$}\label{section:rpi}

  R-$\pi$ is more intense relative to the experimental system. The difference in intensity
  ranges from 3 times too high in the ordered parallel displaced system to 15 times too 
  high in the ordered sandwiched configuration (Table~\ref{table:relative_inensities_300K}).

  The peak shape of the $q_r$ cross-section of R-$\pi$ is qualitatively different
  from experiment (Figures~\ref{fig:exp_rsection_fit} and~\ref{fig:sim_rsection_fit}).
  We use simulated XRD generated from the ordered parallel displaced configuration
  as an example in Figure~\ref{fig:rpi_exp_comparison}. The $q_z$ cross-section 
  fits a Lorentzian profile relatively well, although its full width at half maximum
  (FWHM) is 48 \% broader than the experimental cross-section 
  (Figures~\ref{fig:exp_zsection_fit} and ~\ref{fig:sim_zsection_fit}). We attempted
  to fit a Lorentzian profile to the $q_r$ cross-section and although the FWHM agrees
  with experiment within error, the fit is not optimal due to the three sharp peaks
  that appear near $q_r$ = 0.

  An exact quantitative comparison between experiment and simulation is not feasible. 
  Experimental peaks broaden due to effects that we can not easily simulate such
  as finite size crystalline domains and instrumental resolution. In our simulated
  XRD patterns, the system size imposes limitations on the resolution, which makes
  it difficult to reliably fit peaks. This is especially problematic when comparing
  the FWHM of the $q_z$ cross-section of R-$\pi$, where the experimental FWHM 
  (0.11 $\AA^{-1}$) is similar to the bin size in the $q_z$ direction (ca. 0.07 $\AA^{-1}$).
  Additionally, we model each atom with a Gaussian sphere of electron density. 
  The true electronic structure is likely much more complicated. The subsequent
  discussion primarily serves to demonstrate how we can qualitatively change the
  peak shapes to be in better agreement with experiment.

% BJC9: old version of above paragraph
%  The full width at half maximums (FWHM) of the $q_r$ and $q_z$ cross-sections
%  of R-$\pi$ are broader than experiment (Figure~\ref{fig:rpi_exp_comparison}).
%  The FWHM of the $q_z$ cross-section is 76 \% larger than experiment and the
%  $q_r$ cross-section FWHM is 21\% larger than experiment. Note that the
%  experimental data is best fit with a Lorentzian function while the simulated
%  patterns are best fit using a Gaussian function. The simulated spectra do not
%  incorporate the effects of finite domain sizes, pressure broadening and
%  instrumental broadening which would alter the peak shape. In the absence of
%  these and other peak broadening effects, the experimental peak would actually
%  be narrower, further increasing the discrepancy with simulation.

@@

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{exp_rsection_fit.png}
  \caption{FWHM = 0.792 $\pm$ 0.009 $\AA^{-1}$}\label{fig:exp_rsection_fit}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{sim_rsection_fit.png}
  \caption{FWHM = 0.796 $\pm$ 0.011 $\AA^{-1}$}\label{fig:sim_rsection_fit}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{exp_zsection_fit.png}
  \caption{FWHM = 0.110 $\pm$ 0.003 $\AA^{-1}$}\label{fig:exp_zsection_fit}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{sim_zsection_fit.png}
  \caption{FWHM = 0.163 $\pm$ 0.011 $\AA^{-1}$}\label{fig:sim_zsection_fit}
  \end{subfigure}
  \caption{
%  There are clear qualitative differences between the experimental WAXS pattern 
%  ((a) and (c)) and the simulated XRD pattern resulting from simulations 
%  of the ordered parallel displaced configuration ((b) and (d)). 
  We compared the experimental WAXS pattern ((a) and (c)) to the simulated XRD
  pattern resulting from simulations of the ordered parallel displaced 
  configuration ((b) and (d)). The $q_r$ cross-section of R-$\pi$ ((a) and (b))
  is qualitatively different between experiment (a) and simulation (b). We fit
  Lorentzian profiles to each peak and the FWHM of the simulated pattern agrees
  with experiment within error, however the fit to the simulated data is 
  obscured by the three sharp peaks which appear near $q_r$ = 0. The $q_z$ 
  cross-sections of R-$\pi$ ((c) and (d)) are qualitatively similar. Each fits
  a Lorentzian profile well, however the FWHM of the simulated cross-section (d)
  is 48 \% larger than the experimental cross-section.}\label{fig:rpi_exp_comparison}
  \end{figure}

  We have identified a number of potential ways that simulation conditions,
  forcefield and initial configurations could affect the intensity and shape of
  R-$\pi$. We assessed each using a simplified model. Our simplified model 
  epresents each monomer head group as a single point and places them in either
  a parallel displaced or sandwiched configuration. We chose to explore the 
  influence of thermal noise in each dimension, the effect of $z$-direction 
  distance correlations between scatterers within columns, as well as correlation
  between columns since these are the variables we believe have the largest 
  influence on peak shape and intensity.

  As a baseline, we created a simplified model using the same amount of
  disorder present in our atomistic simulations. There are two sources of
  disorder to consider: thermal motion of atoms and quenched disorder which is
  created by rapid structural rearrangement during early equilibration and 
  is largely locked into place for the remainder of the simulation. The latter
  has important implications for the reproducibility of this work since both 
  initial configuration and initial velocity randomization can lead to 
  different quenched configurations. We measured thermal disorder by
  calculating the standard deviation of the distribution of head group center 
  of mass positions from their average positions. We measured quenched 
  disorder by calculating the standard deviation of the distribution of head
  group center of masses from their idealized average positions. In the 
  $z$-direction, we measured the deviation of the head group center of masses
  from an equally spaced column of head groups. In the $xy$ plane, we 
  calculated the deviation in radial position from the pore center and the
  angular deviation from equally spaced points surrounding the pore center.
  This method for calculating quenched order inherently includes thermal 
  disorder. The values measured for each system and type of disorder are given
  in Table~\ref{table:quenched_disorder}.

  The expected maximum intensity of R-$\pi$ can be calculated from a single
  trajectory of uncorrelated configurations rather than from the average of 
  multiple trajectories of correlated frames. Theoretically, in order to
  mimic the simulated system, we should build an initial configuration which 
  has the same amount of quenched disorder as our simulations and then create
  a trajectory to which we add thermal noise to each subsequent frame. In this
  scenario, the positions of each scatterer will not change in a way that 
  significantly affects the intensity of R-$\pi$ since most of its character 
  is defined by the far more disordered initial configuration. To uncover the
  expected intensity of R-$\pi$, we would need to average over many such 
  trajectories. To bypass this expensive calculation, we instead calculate the
  R-$\pi$ intensity from a trajectory consisting of many independent initial 
  configurations. This simulates the long time scales over which we expect 
  columns to sample many configurational states. %We use the latter method 
  %for the remainder of this section since we can extract the trends we 
  %need using a single trajectory at a fraction of the computational cost.

  When we simulated the diffraction pattern of a parallel displaced simple
  system trajectory of 1000 independent configurations, the
  $q_r$ and $q_z$ profiles of R-$\pi$ are characterized by sharp peaks. We made
  this simple system, and all to follow, 4 times taller in the $z$-direction in
  order to access higher $q_z$ resolution. Figure~\ref{fig:sf_qy_nocorrelation}
  and~\ref{fig:sf_qz_correlation} shows the $q_y$ and $q_z$ cross-sections of
  R-$\pi$ generated by simulating the structure factor of the trajectory. The
  distance between peaks in the $q_y$ cross-section is equivalent to the distance
  between pores in real space.

@@

  The distances between scatterers within each column are correlated with each
  other. In the simplified system described above, we added Gaussian noise to
  the mean positions of an equally spaced column of points. This ignores any
  interaction head groups might have with each other. To remedy this, we drew
  random samples from a multivariate normal distribution defined by the mean
  positions of an equally spaced column of points. We gave the distribution at
  each point along the column the same standard deviation, $\sigma_z$, shown in
  Table~\ref{table:quenched_disorder}. We defined a covariance matrix such
  that the covariance, $v$, of the distance, $d$, between scatterers decays
  exponentially from $v$ according to the equation $ve^{-d/L}$, where $L$ is the
  correlation length. Experimentally, the correlation length can be estimated as
  the inverse of the FWHM of the $q_z$ cross-section of R-$\pi$. Accordingly,
  the correlation length of the experimental system is $\geq$ 9 $\AA$. The value
  may be larger since the peak is likely broadened. We simulated the structure
  factor of a simple system built as described above with a correlation length of
  30 $\AA$. We chose 30 $\AA$ in order to exaggerate the effect of correlation
  length for visual purposes. Subsequent structure factor calculations are all
  done with a correlation length of 9 $\AA$. The shape of the $q_y$ cross-section
  (Figure~\ref{fig:sf_qy_correlation}) remains the same while the the $q_z$
  cross-section (Figure~\ref{fig:sf_qz_correlation}) broadens, as seen in
  experiment. Additionally, there is a near 20 \% increase in the maximum
  intensity of R-$\pi$ when the distances between scatterers are correlated.

>>>>

  We further refined our structural understanding of the system by simulating X-ray 
  diffraction patterns produced from equilibrated MD trajectories and comparing them
  to experiment. We tested systems built with 5 columns per pore in the parallel 
  displaced and sandwiched configurations at 300 K in the ordered and disordered
  basins. We generated simulated patterns using the equilibrated portion of each 
  simulation trajectory. The patterns for all structures are shown and compared to
  experiment in Figure~\ref{fig:XRDsim}.
  %MRS10: Would be btter to lead with something describing what we found, not a caveat, just for clarity. Might belong better in the previous paragraph if left here, since it is a clarifying point that should be communicated, not an actual finding.
  The diffraction patterns show some noise, especially along the axis where $q_r$=0. There are
  two main reasons for this noise. First is that we angle averaged the 3D structure factor 
  (see section~\ref{method:xrd}) which inherently gives less data as $q_r$ approaches 0. Second, 
  our simulations are not long enough to sample enough truly independent configurations within
  their respective metastable basins. By sampling more uncorrelated equilibrium configurations,
  we can suppress some of the noise (See section S-\ref{S-section:xrd_noise}).

@@

  The simulated XRD patterns show moderate qualitative agreement with experiment. R-alkanes and
  R-pores appear in the expected location. R-pores is more intense than experiment, likely
  because we are simulating a near perfect, infinite hexagonal array. The real system has 
  defects and domain misalignment which decreases the overall intensity.
  %MRS10: would domain misalignment make a difference in itensity?  I though we decided it wouldnt (it would more broaden?).
  %You can talk about how you WILL analyze the intensity diffence later on, that it doesn't need to be analyzed here.
  % MRS4: comments from Matt suggest that domain misalignment shouldn't alone be responsible for 
  % the difference in intensity ratio. Some of the experiments should help explain it better. 	
  % Let's talk!
  % BJC4: So a combo of: domain misalignment, defects, slow dynamics of simulated system, only 4 
  % pores in unit cell, tortuosity (?)
  % MRS5: maybe, there's a question of how we can demonstrate the magnitude of any of these 
  % effects.
  R-spots appears to be weaker, except for the ordered basin sandwiched configuration, in the 
  simulated patterns. R-spots is also partially engulfed by the wide R-$\pi$ reflection. R-$\pi$
  appears at a lower $q_z$ value than experiment. In all systems, the reflection reaches its 
  maximum at $q_z < 1.5~\AA^{-1}$ which means that monomers prefer to stack at least 4.2 \AA~apart
  rather than 3.7 \AA. R-double does not appear in any of the patterns.

@@

  We quantified the numerical discrepancies present when comparing the relative intensities
  of each reflection of interest between experimental and simulated patterns. 
  Table~\ref{table:relative_inensities_300K} shows the relative intensities of each reflection for 
  all systems tested. The patterns are normalized so that the average intensity of R-alkanes 
  must equal 1. We measured the approximate intensity of R-$\pi$ and R-double by measuring the
  intensity of the appropriate peaks of the cross-section of the 2D pattern at $q_r=0$. The 
  relative intensity of R-$\pi$ is significantly higher than experiment in our simulations. 
  %MRS10: weird transition between these two; it's a result that interrupts the discussion of methods.
  R-spots is measured as the average intensity within the region bounded by a 'spot'. We 
  identified spots based on visual inspection. 
  %MRS10: what criteria were used?  Need to provide what was used to make the decision.
If the spots were not easily discernible, then the 
  intensity was taken as that of the intersection of R-alkanes at half the $q_z$ value of
  R-$\pi$, as that is where it appears experimentally. The intensity of R-spots is slightly
  lower than experiment in all cases except for the ordered basin sandwiched configuration which 
  agrees with experiment. There is no R-double intensity to be measured. %MRS10: specific where no R-double intensity was measured. None of the simulations, correct?
Figures are provided in section S-\ref{S-section:xrd_intensities} to further clarify these measurements. %MRS10: Specify what figures clarify what things.  Also, use active voice in this section.

  \begin{table}[h]
  \centering
  \begin{tabular}{c|ccccc}
  \toprule
 		   & \multicolumn{5}{c}{Configuration} \\
  \hline
             &            &            & Parallel  & Disordered & Disordered         \\
  Reflection & Experiment & Sandwiched & Displaced & Sandwiched & Parallel Displaced \\
  \midrule
  R-alkanes  & 1.0        &  1.0       &  1.0      & 1.0        & 1.0                \\  
  R-spots    & 1.3        &  1.3       &  1.2      & 1.1        & 1.2                \\
  R-$\pi$    & 2.8        & 44.0       &  7.7      & 8.4        & 10.1               \\
  R-double   & 0.9        &  --        & --        &  --        & --                 \\ 
  \bottomrule
  \end{tabular}
  \caption{The simulated XRD patterns of the systems tested, normalized so that the average 
  intensity of R-alkanes equals 1, show R-$\pi$ reflections that are significantly higher 
  than experiment and R-spots reflections that are slightly lower than experiment. R-double
  does not appear in any patterns, and thus has no measurable intensity.}
  \label{table:relative_inensities_300K} 
  \end{table}
  %BJC5: Bold font is caused by achemso document type
  %MRS7: huh.
  %BJC6: yea, specifically the one for j phys chem B. I compiled with a different document type and the captions looked normal
  There are clear differences between simulated and experimental results that must be addressed
  and justified. We will explore the possible origin of select experimental reflections and form 
  structural hypotheses related to their appearance in our simulated patterns. Specifically, we 
  want answer:  
  %MRS10: should the ``why is R-pi in not quite the right place'' be here, to have those sorts of questions more centralized?
  \begin{enumerate}
	\item What is the origin of R-spots?
  	\item Why is the intensity of R-$\pi$ so much greater than experiment?
  	\item What is the origin of R-double?
  \end{enumerate}

@@

  \subsubsection{The origin of R-spots}\label{section:rspots}
%MRS10: I rephrased the below, check the diffs.
  In order to more clearly study R-spots, we equilibrated an ordered basin sandwiched 
  configuration at 280K. Many force fields do not obtain the correct equilibrium structure at precisely the experimental temperature. Wang et al. highlighted the 
  shortcoming of some of the most popular protein force field in predicting the temperature
  dependence of protein structural ensembles \cite{wang_building_2017}. It is therefore possible that   % page 12
  our membrane system, simulated at 300K, is overly structured compared to the experimental structure at 300K.
  We found that lowering the temperature better resolved the R-spots (Figure~\ref{fig:sandwiched280K}), suggesting that our 'effective' temperature might indeed be too high, overstabilizing structure. 
%  The relative intensity 
%  of R-spots in this system is slightly higher than experiment (1.5 vs. 1.3)

@@

  Evidence 
%MRS10: added
from the simulations
  strongly suggests that the R-spots signal is not a result of alkane
  chain tilt. Previous literature has attributed the spots in this particular
%MRS10: clarified that it was THIS exact data, and not similiar images. Correct if wrong.
WAXS 
%pattern 
dataset
as the product of tilted alkane chains~\cite{feng_scalable_2014}.
  We measured the tilt angle of the alkane chains of the 280K system and showed
  that it equilibrates to an average tilt angle of
  -2\degree~(Fig.~\ref{fig:tilt}), far from the 37\degree~tilt angle previously
  used to explain R-spots. 
  %MRS8: did you do a few different things here as sensitivity analysis to show that the finding was robust?  If so, putting some of that in would be useful.
  %BJC7: make it clear that it is the 280K. Reference all configurations where tilt angles were calculated.

@@

  %BJC9: I think Matt gets confused here. I put a more detailed explanation in 
  % the supplemental information. Do you think that's enough, or should I add
  % to this figure?
  %MRS10: rephrased.
  The evidence strongly suggests that R-spot pattern occurs because tails pack together hexagonally. 
  We plotted the center of masses of the 
  first four tail atoms of each tail originating from each head group 
  (Figure \ref{fig:centroids}). We measured the angle between each center of
  mass and its nearest neighbor center of masses with respect to the $xy$ 
  plane of the membrane. We see distinct peaks in the distribution of these 
  angles located ca. $-60\degree$, $0\degree$, and $60\degree$ which is 
  consistent with a hexagonally packed configuration 
  (Figure~\ref{fig:layered_tails}). 
  %MRS10: is it importat that the ordering dies off? IF there is ordering, it will show in the spectra, and it doesn't matter that it dies off in other parts of the spectra.
This ordering dies off at the tail ends,
  where there is more space to fill causing them to pack nearly isotropically.
  See Section~S-\ref{S-tail_packing} for a more detailed explanation of the 
  calculation and packing distributions generated from different sections of
  the tails.

  The peaks in the nearest neighbor angle distribution due to the hexagonal packing are consistent with the
  location of R-spots. The 2D Fourier transform of a simple hexagonal array 
  constructed based on the peak angles in Figure~\ref{fig:layered_tails}
  shows reflections in the same locations as R-spots, in addition to vertical 
  stacking reflections along the $q_z$ axis that would intersect with R-$\pi$
  (Figure~\ref{fig:hexagonal_ft}).

@@

  \subsubsection{Discrepancies between experimental and simulated R-$\pi$}\label{section:rpi}
  %MRS10: in the list of questions above, you just talk about intensity, but this title talks about more discrepnacies. It could go either way (as I mention before, there's the issue of R-pi placement, too), but it should be consistent.
  The experimentally measured R-$\pi$ is more intense relative to the experimental system. The difference in intensity
  ranges from 3 times too high in the ordered parallel displaced system to 15 times too 
  high in the ordered sandwiched configuration (Table~\ref{table:relative_inensities_300K}).
  %MRS10: should specify this is in the 300 K simulations.

  The peak shape of the $q_r$ cross-section of R-$\pi$ in the simulated XRD is qualitatively different
  from experiment (Figures~\ref{fig:exp_rsection_fit} and~\ref{fig:sim_rsection_fit}).
  We use simulated XRD generated from the ordered parallel displaced configuration
  as an example in Figure~\ref{fig:rpi_exp_comparison}. The $q_z$ cross-section 
  fits a Lorentzian profile relatively well, although its full width at half maximum
  (FWHM) is 48\% broader than the experimental cross-section 
  (Figures~\ref{fig:exp_zsection_fit} and ~\ref{fig:sim_zsection_fit}). We attempted
  to fit a Lorentzian profile to the $q_r$ cross-section and although the FWHM agrees
  with experiment within error, the fit is not optimal due to the three sharp peaks
  that appear near $q_r$ = 0.

  An exact quantitative comparison between experiment and simulation is not feasible for several reasons. 
  Experimental peaks broaden due to effects that we can not easily simulate such
  as finite size crystalline domains and instrumental resolution. In our simulated
  XRD patterns, the system size imposes limitations on the resolution, which makes
  it difficult to reliably fit peaks. This is especially problematic when comparing
  the FWHM of the $q_z$ cross-section of R-$\pi$, where the experimental FWHM 
  (0.11 $\AA^{-1}$) is similar to the bin size in the $q_z$ direction (ca. 0.07 $\AA^{-1}$).
  Additionally, we model each atom with a Gaussian sphere of electron density. 
  The true electronic structure is likely much more complicated. 
  %MRS10: modified
  However, we can explore reasons for which differences in the simulation structure could changes the peak shapes to be in better agreement with experiment.

% BJC9: old version of above paragraph
%  The full width at half maximums (FWHM) of the $q_r$ and $q_z$ cross-sections
%  of R-$\pi$ are broader than experiment (Figure~\ref{fig:rpi_exp_comparison}).
%  The FWHM of the $q_z$ cross-section is 76\% larger than experiment and the
%  $q_r$ cross-section FWHM is 21\% larger than experiment. Note that the
%  experimental data is best fit with a Lorentzian function while the simulated
%  patterns are best fit using a Gaussian function. The simulated spectra do not
%  incorporate the effects of finite domain sizes, pressure broadening and
%  instrumental broadening which would alter the peak shape. In the absence of
%  these and other peak broadening effects, the experimental peak would actually
%  be narrower, further increasing the discrepancy with simulation.

@@

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{exp_rsection_fit.png}
  \caption{FWHM = 0.792 $\pm$ 0.009 $\AA^{-1}$}\label{fig:exp_rsection_fit}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{sim_rsection_fit.png}
  \caption{FWHM = 0.796 $\pm$ 0.011 $\AA^{-1}$}\label{fig:sim_rsection_fit}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{exp_zsection_fit.png}
  \caption{FWHM = 0.110 $\pm$ 0.003 $\AA^{-1}$}\label{fig:exp_zsection_fit}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{sim_zsection_fit.png}
  \caption{FWHM = 0.163 $\pm$ 0.011 $\AA^{-1}$}\label{fig:sim_zsection_fit}
  \end{subfigure}
  \caption{
%  There are clear qualitative differences between the experimental WAXS pattern 
%  ((a) and (c)) and the simulated XRD pattern resulting from simulations 
%  of the ordered parallel displaced configuration ((b) and (d)). 
    %MRS10: don't make the thesis what was done, but what the reader should know. 
  We compared the experimental WAXS pattern ((a) and (c)) to the simulated XRD
  pattern resulting from simulations of the ordered parallel displaced 
  configuration ((b) and (d)). The $q_r$ cross-section of R-$\pi$ ((a) and (b))
  is qualitatively different between experiment (a) and simulation (b). We fit
  Lorentzian profiles to each peak and the FWHM of the simulated pattern agrees
  with experiment within error, however the fit to the simulated data is 
  obscured by the three sharp peaks which appear near $q_r$ = 0. The $q_z$ 
  cross-sections of R-$\pi$ ((c) and (d)) are qualitatively similar. Each fits
  a Lorentzian profile well, however the FWHM of the simulated cross-section (d)
  is 48\% larger than the experimental cross-section.}\label{fig:rpi_exp_comparison}
  \end{figure}

  We have identified a number of potential ways that simulation conditions,
  forcefield and initial configurations could affect the intensity and shape of
  R-$\pi$, and assess each using a simplified model of the pore. Our simplified model 
  represents each monomer head group as a single point and places them in either
  a parallel displaced or sandwiched configuration. We chose to explore the 
  influence of thermal noise in each dimension, the effect of $z$-direction 
  distance correlations between scatterers within columns, as well as correlation
  between columns since these are the variables we believe have the largest 
  influence on peak shape and intensity.

  %MRS10: somewhere in here, we need to address whether quenched and static disorder are 
  % the same experimentally. If not the same, then it's glassy, if they are the same, then 
  % we are making the assumption that the the timescales of motion of that noice are > 1 microsecond but < 1 second.
  We created a simplified baseline model using the same amount of
  disorder present in our atomistic simulations. There are two sources of
  disorder to consider: thermal motion of atoms and quenched disorder which is
  created by rapid structural rearrangement during early equilibration and 
  is largely locked into place for the remainder of the simulation. The latter
  has important implications 
  %MRS10: which are the important implications?
for the reproducibility of this work since both 
  initial configuration and initial velocity randomization can lead to 
  different quenched configurations. We measured thermal disorder by
  calculating the standard deviation of the distribution of head group center 
  of mass positions from their average positions. We measured quenched 
  disorder by calculating the standard deviation of the distribution of head
  group center of masses from their idealized average positions. In the 
  $z$-direction, we measured the deviation of the head group center of masses
  from an equally spaced column of head groups. In the $xy$ plane, we 
  calculated the deviation in radial position from the pore center and the
  angular deviation from equally spaced points surrounding the pore center.
  This method for calculating quenched order inherently includes thermal 
  disorder. The values measured for each system and type of disorder are given
  in Table~\ref{table:quenched_disorder}.

  %MRS10: this is more of a methodological thesis than is ideal. Can you make it more about what we are learning, or at least what was done and the reasoning?
  The expected maximum intensity of R-$\pi$ can be calculated from a single
  trajectory of uncorrelated configurations rather than from the average of 
  multiple trajectories of correlated frames. Theoretically, in order to
  mimic the simulated system, we should build an initial configuration which 
  has the same amount of quenched disorder as our simulations and then create
  a trajectory to which we add thermal noise to each subsequent frame. In this
  scenario, the positions of each scatterer will not change in a way that 
  significantly affects the intensity of R-$\pi$ since most of its character 
  is defined by the far more disordered initial configuration. To uncover the
  expected intensity of R-$\pi$, we would need to average over many such 
  trajectories. To bypass this expensive calculation, we instead calculate the
  R-$\pi$ intensity from a trajectory consisting of many independent initial 
  configurations. This simulates the long time scales over which we expect 
  columns to sample many configurational states. %We use the latter method 
  %for the remainder of this section since we can extract the trends we 
  %need using a single trajectory at a fraction of the computational cost.

  %MRS10: again, more of a methodolgical thesis.  Will not be clear to the reader what they are 
  % expecting to learn, which is the difference in the simulation that could explain the differences in the spectra.
  % I'm thinking some of this should be in the supporting info.  Maybe not, if key enough, but it should still be in support of the conclusion. 
  When we simulated the diffraction pattern of a parallel displaced simple
  system trajectory of 1000 independent configurations, the
  $q_r$ and $q_z$ profiles of R-$\pi$ are characterized by sharp peaks. We made
  this simple system, and all to follow, 4 times taller in the $z$-direction in
  order to access higher $q_z$ resolution. Figure~\ref{fig:sf_qy_nocorrelation}
  and~\ref{fig:sf_qz_correlation} shows the $q_y$ and $q_z$ cross-sections of
  R-$\pi$ generated by simulating the structure factor of the trajectory. The
  distance between peaks in the $q_y$ cross-section is equivalent to the distance
  between pores in real space.

@@

  The distances between scatterers within each column are correlated with each
  other. In the simplified system described above, we added Gaussian noise to
  the mean positions of an equally spaced column of points. This ignores any
  interaction head groups might have with each other. To remedy this, we drew
  random samples from a multivariate normal distribution defined by the mean
  positions of an equally spaced column of points. We gave the distribution at
  each point along the column the same standard deviation, $\sigma_z$, shown in
  Table~\ref{table:quenched_disorder}. We defined a covariance matrix such
  that the covariance, $v$, of the distance, $d$, between scatterers decays
  exponentially from $v$ according to the equation $ve^{-d/L}$, where $L$ is the
  correlation length. Experimentally, the correlation length can be estimated as
  the inverse of the FWHM of the $q_z$ cross-section of R-$\pi$. Accordingly,
  the correlation length of the experimental system is $\geq$ 9 $\AA$. The value
  may be larger since the peak is likely broadened. We simulated the structure
  factor of a simple system built as described above with a correlation length of
  30 $\AA$. We chose 30 $\AA$ in order to exaggerate the effect of correlation
  length for visual purposes. Subsequent structure factor calculations are all
  done with a correlation length of 9 $\AA$. The shape of the $q_y$ cross-section
  (Figure~\ref{fig:sf_qy_correlation}) remains the same while the the $q_z$
  cross-section (Figure~\ref{fig:sf_qz_correlation}) broadens, as seen in
  experiment. Additionally, there is a near 20\% increase in the maximum
  intensity of R-$\pi$ when the distances between scatterers are correlated.
